Quantum Walks and Application to Quantum Money\\[1em] \normalsize McMaster University\\ Department of Computing and Software

### 2025-04-10 draft.tex
draft updated

\usepackage{amsmath, amssymb, amsthm}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{physics}
\usepackage{algorithm}
% \usepackage{algorithmic}
\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{setspace}
\usepackage{algorithm}
\usepackage{algpseudocode}

\newcommand{\opnorm}[1]{\left\lVert #1 \right\rVert}

\usepackage[a4paper, left=1.5cm, right=1.5cm, top=2cm, bottom=2cm]{geometry}

% Formatting
\setlength{\parskip}{0.5em}
\setlength{\parindent}{0pt}
\singlespacing

@@

\title{
  Quantum Walks and Application to Quantum Money\\[1em]
  \large Supervised by Dr. Jake Doliskani\\[3em]
  \normalsize McMaster University\\
  Department of Computing and Software
}
\author{Ali Mousavi}
\date{April 2025}

@@

\newpage
\textbf{\large Abstract}

@@

\medskip\noindent\textbf{This Work – Motivation and Contributions.} 
In this thesis, we adapt Zhandry’s group-action quantum money scheme by replacing its use of the Quantum Fourier Transform with the \emph{Quantum Hartley Transform} (QHT) over finite abelian groups. The motivation behind this substitution is multifold. First, using the Hartley transform causes the banknotes to have \emph{real} amplitudes instead of complex amplitudes. We believe this shift from complex to real quantum states may offer both computational and theoretical advantages. For example, certain mathematical identities hold for any two real orthonormal bases that do not generally hold for complex bases; such differences can affect the properties of quantum states and were a barrier in prior analyses (indeed, an attempted “quantum lightning” construction in [31] failed due to properties of complex phases that do not arise with real amplitudes). Beyond this specific context, our work is a first step toward employing real-valued quantum transforms in quantum cryptography and quantum algorithms. To the best of our knowledge, this is the first significant quantum cryptographic construction that makes essential use of the quantum Hartley transform. We hope that exploring real-amplitude quantum states will inspire further research into new quantum algorithms and optimizations for real-valued transforms.

@@

\medskip\noindent\textbf{Thesis Organization.} 
The remainder of this thesis is organized as follows. In **Chapter 2**, we provide background on the basics of quantum computation and the theory of group actions, including the concept of cryptographic group actions which underlie our money scheme. **Chapter 3** covers quantum walks, both continuous-time and discrete-time, outlining their differences and importance in quantum algorithms. In **Chapter 4**, we apply the quantum walk framework to group actions: we describe how a continuous-time quantum walk on a group action can be simulated efficiently, a crucial step for our serial number extraction algorithm. **Chapter 5** details the quantum money scheme itself and the role of the Hartley transform in its construction. We explain how the scheme is formulated and discuss the advantages and challenges introduced by using the Hartley transform. Finally, **Chapter 6** presents the verification algorithm for the Hartley-based quantum money. This chapter shows how we use quantum walks (as developed in Chapter 4) to reliably verify banknotes and compute their serial numbers, thereby addressing the verification issues that arose from the use of real amplitudes.
\vspace{1em}

\chapter{Background of Quantum Computation and Group Actions}

@@

One example important to this work is the \textbf{quantum Fourier transform} (QFT). For an abelian group $G$ of order $N$, the QFT is the unitary map
\[
|g\rangle \mapsto \frac{1}{\sqrt{N}}\sum_{h \in G} \chi(h,g)\, |h\rangle,
\]
where $\{\chi(h,.)\}$ are the characters (complex exponential homomorphisms) of $G$. When $G = \mathbb{Z}_N$ (the integers mod $N$), the QFT takes $|j\rangle$ to $\frac{1}{\sqrt{N}}\sum_{k=0}^{N-1} e^{2\pi i j k / N} |k\rangle$.

Let $G \cong \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$ be a decomposition of $G$. The character $\chi(a, x)$ is then given by
\[
\chi(a, x) = \omega_{N_1}^{a_1x_1} \cdots \omega_{N_k}^{a_kx_k}, \quad \text{where } \omega_M = \exp(2\pi i/ M).
\]

@@

\paragraph{The cmpIndex Algorithm.}
Given a state $\ket{G^{(h)} * x}$, one can recover $h$ using phase kickback. Starting from $\ket{G^{(h)} * x} \ket{0}$, apply:
\begin{itemize}
    \item QFT to second register,
    \item Controlled unitary $\sum_{k \in G} U_k \otimes \ket{k}\bra{k}$,
    \item Inverse QFT to second register.
\end{itemize}
This yields $\ket{G^{(h)} * x} \ket{h}$, extracting $h$ efficiently.

@@

\paragraph{Definition.}
For $G = \mathbb{Z}_N$, define
\[
\operatorname{QHT}_N: \ket{a} \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\left(\frac{2\pi a y}{N}\right) \ket{y}, \quad \text{where } \cas(x) = \cos(x) + \sin(x).
\]

For general abelian $G = \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$, the Hartley transform of $f : G \to \mathbb{R}$ is
\begin{equation}
    \operatorname{QHT}_G(f)({a}) = \frac{1}{\sqrt{|G|}} \sum_{{y} \in G} \cas(2\pi \langle {a}, \alpha({y}) \rangle) f({y}),
\end{equation}
where $\alpha({y}) = y_1 / N_1 + \cdots + y_k / N_k$.

\paragraph{Unitary Equivalence.}
Using the identity
\[
\cas(2\pi \langle {a}, \alpha({y}) \rangle) = \frac{1 - i}{2}\chi({a}, {y}) + \frac{1 + i}{2}\chi(-{a}, {y}),
\]
we obtain:
\begin{equation}
\operatorname{QHT}_G = \frac{1 - i}{2} \operatorname{QFT}_G + \frac{1 + i}{2} \operatorname{QFT}_G^*, \label{eq:QHT_QFT}
\end{equation}
showing that $\operatorname{QHT}_G^2 = {1}$ and hence the QHT is unitary.

\paragraph{Quantum Formulation.}
The quantum Hartley transform over $G$ maps:
\[
\sum_{{x} \in G} f({x}) \ket{{x}} \mapsto \sum_{{a} \in G} \operatorname{QHT}_G(f)({a}) \ket{{a}}.
\]
Because the cas function is real-valued, the resulting amplitudes remain real—a fact that plays a crucial role in real-amplitude quantum protocols.

@@

\chapter{Quantum Walks -- Continuous and Discrete}

Quantum walks are the quantum-mechanical counterparts of classical random walks on graphs. Instead of a probability distribution over vertices, a quantum walk evolves a superposition of vertex states. Interference between multiple paths leads to markedly different behavior from classical walks, often yielding faster spreading and algorithmic speedups~\cite{physics_stackexchange}.

In particular, both major models of quantum walks --- continuous-time and discrete-time --- achieve \emph{ballistic spreading} (standard deviation growing linearly in time) in contrast to the diffusive $O(\sqrt{t})$ spread of classical random walks~\cite{childs2003quantum}. Quantum walks have become a central framework in quantum computing, underpinning various algorithms and even universal quantum computation~\cite{file_gwrqpr3gkui6gav6sze4dj}.

@@

Continuous-time quantum walks have been employed to achieve algorithmic speedups. A famous example is the \emph{glued trees traversal problem} by Childs et al. (2003), where a CTQW on a graph formed by two randomly "glued" binary trees can find the unique exit vertex exponentially faster than any classical algorithm. This result gave the first exponential separation between quantum and classical query complexity in a black-box setting.

Another application is \emph{spatial search} on graphs. Childs and Goldstone (2004) showed that a CTQW can find a marked vertex on a $d$-dimensional lattice faster than classical diffusion. In the limit $d \to \infty$, the CTQW dynamics converge to the continuous Dirac equation.

@@

In 2004, Szegedy proposed a different formulation of DTQW based on classical Markov chains. Here, no coin register is needed. Instead, the walk is defined on $\mathbb{C}^N \otimes \mathbb{C}^N$, with basis states $\ket{j,k}$ corresponding to directed edges.

@@

Both DTQW and CTQW have been used to achieve quadratic speedups in search problems. In many cases, the two models can simulate each other. DTQWs can even achieve \textbf{universal quantum computation}: Childs et al. (2009) constructed a coined DTQW that simulates an arbitrary quantum circuit.

@@

    \item \textbf{Infinite Tree (Free Group):} The free group on two generators $a$ and $b$ has a Cayley graph that is an infinite 4-regular tree. Each vertex corresponds to a word in $\{a^{\pm1}, b^{\pm1}\}$, and edges represent multiplication by $a$ or $b$. A finite portion of this graph is shown in Figure~\ref{fig:cayley-free-group}.

@@

To see this, we have
\[
        A\ket{\hat{a}} = A.\frac{1}{\sqrt{|G|}}\sum_{y \in G}\chi(a,y)\ket{y} = \frac{1}{\sqrt{|G|}}\sum_{y \in G}\chi(a,y).A\ket{y}
        \]
        \[
        = \frac{1}{\sqrt{|G|}}\sum_{y \in G}\chi(a,y).\sum_{q \in Q}\ket{qy}
        \]
        Consider $\beta = qy$. Then:
        \[
         = \frac{1}{\sqrt{|G|}}\sum_{q \in Q}\chi(a,q)\sum_{\beta}\chi(a,\beta)\ket{\beta} = \sum_{q \in Q}\chi(a,q) . \ket{\hat{a}}
        \]
        \[
         = \lambda_a\ket{\hat{a}}
        \]

@@

In particular, for a continuous-time quantum walk starting at the identity element:
\[
\ket{\psi(t)} = e^{-iAt} \ket{e},
\]
the resulting state remains a \emph{class function}, i.e., the amplitude $\braket{g|\psi(t)}$ depends only on structural properties of $g$ (such as Hamming weight or conjugacy class).

@@

\[
    A\ket{G^{(h)}*x} = \frac{1}{\sqrt{|G|}}.\sum_{g \in G}\chi(g,h) A.\ket{g*x}
    \]

    \[
    = A\ket{G^{(h)}*x} = \frac{1}{\sqrt{|G|}}.\sum_{g \in G}\chi(g,h)\sum_{q \in Q}\ket{q*(g*x)}
    \]
    Consider $\beta = qg$. Then:
    \[
    = = \frac{1}{\sqrt{|G|}}\sum_{q \in Q}\sum_{\beta \in G}\chi(\beta,h)\chi(q,h)\ket{\beta * x}
    \]

    \[
    = \sum_{q \in Q}\chi(q,h).\ket{G^{(h)}*x} = \lambda_h \ket{G^{(h)}*x}
    \]

@@

\chapter{Simulating Group Action Quantum Walks}

From this point onward, we assume that the generating set \( Q \) is small, specifically \( \abs{Q} = \operatorname{poly}(\log \abs{G}) \). In the first construction, where the vertex set is \( X = G \), the continuous-time walk \( e^{-iAt} \) can be implemented efficiently even for values of \( t \) that are exponentially large in \( \log \abs{G} \). This efficiency stems from the identity:
\begin{equation}
    \label{eq:adjmat-fb}
    e^{-iAt} = \operatorname{QFT}_G \left( \sum_{a \in G} e^{-i\lambda_a t} \ket{a} \bra{a} \right) \operatorname{QFT}_G^*,
\end{equation}
where both \( \operatorname{QFT}_G \) and its inverse can be implemented in \( \operatorname{poly}(\log \abs{G}) \) time. Moreover, the diagonal unitary \( \ket{a} \mapsto e^{-i\lambda_a t} \ket{a} \) is efficiently realizable since the eigenvalues \( \lambda_a \) can be computed classically to arbitrary precision in polynomial time with respect to \( \log \abs{G} \).

@@

Based on these states, we define the isometry \( T : \mathbb{C}^{2N} \to \mathbb{C}^{2N} \otimes \mathbb{C}^{2N} \) as follows:
\begin{equation}
    \label{eq:walk-isom}
    T := \sum_{j = 0}^{N - 1} \sum_{b \in \{0, 1\}} (\ket{j}\bra{j} \otimes \ket{b}\bra{b}) \otimes \ket{\phi_{jb}}.
\end{equation}

The discrete-time quantum walk operator is then defined as
\[
W = iS(2TT^* - \mathbf{1}),
\]
where \( S \) is the swap operator acting as
\[
S \ket{j_1} \ket{b_1} \ket{j_2} \ket{b_2} = \ket{j_2} \ket{b_2} \ket{j_1} \ket{b_1},
\]
for all \( 0 \le j_1, j_2 < N \) and \( b_1, b_2 \in \{0, 1\} \).

@@

\section{Simulating group action quantum walks.}
Assume \( t = \operatorname{poly}(\log \abs{G}) \). We show that the continuous-time walk \( W = e^{-iAt} \) can be efficiently simulated using the discrete-time quantum walk framework developed in~\cite{childs2010relationship}. Recall the isometry \( T \) defined in Equation~\eqref{eq:walk-isom}, where the states \( \ket{\phi_b} \) are as given in Equation~\eqref{eq:walk-sim}. To carry out the simulation, it suffices to show that the isometry \( T \), its adjoint \( T^* \), and the discrete-time walk operator
\[
W = iS(2TT^* - \mathbf{1})
\]
can all be efficiently implemented in the group-action setting.

@@

Because \( Q \subset G \) is a small set (i.e., \( \abs{Q} = \operatorname{poly}(\log \abs{G}) \)), we can construct an efficient unitary \( V_Q : \mathbb{C}^Q \to \mathbb{C}^Q \) such that
\[
V_Q \ket{0} = \frac{1}{\sqrt{\abs{Q}}} \sum_{q \in Q} \ket{q}.
\]
Applying \( V_Q \otimes \mathbf{1} \) to the state \( \ket{0} \ket{y, 0} \ket{0, 0} \) produces:
\[
\frac{1}{\sqrt{\abs{Q}}} \sum_{q \in Q} \ket{q} \ket{y, 0} \ket{0, 0}.
\]

@@

To implement \( U_0^* : \ket{0} \ket{y, 0} \ket{\phi_{y0}} \mapsto \ket{0} \ket{y, 0} \ket{0, 0} \), we sequentially apply \( V_2^* \), followed by \( V_1^* \), and finally \( V_Q^* \otimes \mathbf{1} \). Since each of these unitaries can be implemented efficiently, it follows that \( U_0^* \) is also efficient.

@@

Next, we demonstrate that the walk unitary
\[
W = iS(2TT^* - \mathbf{1})
\]
can be applied efficiently. This reduces to implementing the reflection \( 2TT^* - \mathbf{1} \), which we analyze as follows:
\begin{align*}
2\ket{0}\bra{0} \otimes TT^* - \mathbf{1}
&= 2\ket{0}\bra{0} \otimes \sum_{y \in X} \sum_{b \in \{0,1\}} \ket{y,b} \ket{\phi_{yb}} \bra{y,b} \bra{\phi_{yb}} - \mathbf{1} \\
&= 2\sum_{y \in X} \sum_{b \in \{0,1\}} \ket{0} \ket{y,b} \ket{\phi_{yb}} \bra{0} \bra{y,b} \bra{\phi_{yb}} - \mathbf{1} \\
&= U_T \left( 2 \sum_{y \in X} \sum_{b \in \{0,1\}} \ket{0} \ket{y,b} \ket{0,0} \bra{0} \bra{y,b} \bra{0,0} - \mathbf{1} \right) U_T^* \\
&= U_T \left( 2\ket{0}\bra{0} \otimes \mathbf{1}_{X,b} \otimes \ket{0,0}\bra{0,0} - \mathbf{1} \right) U_T^*.
\end{align*}

Since \( U_T \), \( U_T^* \), and the operator \( 2\ket{0}\bra{0} \otimes \mathbf{1}_{X,b} \otimes \ket{0,0}\bra{0,0} - \mathbf{1} \) can all be implemented efficiently, it follows that the reflection \( 2\ket{0}\bra{0} \otimes TT^* - \mathbf{1} \) is also efficient.

Finally, for any input state \( \ket{\psi} \), the reflection acts as:
\[
\left(2\ket{0}\bra{0} \otimes TT^* - \mathbf{1} \right) \ket{0} \ket{\psi} = \ket{0} \left(2TT^* - \mathbf{1} \right) \ket{\psi},
\]
confirming that the full walk operator \( W \) can be implemented efficiently in the group-action setting.

@@

\chapter{Quantum Money and Hartley Transform}

@@

\section{Fourier-Based Quantum Money from Group Actions}
Zhandry’s quantum money scheme [31] is built on an abelian group action $G \curvearrowright X$. We have a family of group actions parameterized by the security parameter $n$ (so the size of $G$ grows with $n$). For simplicity, assume $|G| = N$. The scheme also fixes a particular element $x_0 \in X$ that everyone knows (a public base point in $X$).

@@

To address this, our verification algorithm takes a more sophisticated approach:
\begin{enumerate}
    \item Given $(h, |\psi\rangle)$, it first uses the \textbf{quantum walk-based phase estimation} method (to be detailed in Chapter 6) to \emph{compute the eigenphase corresponding to $h$}. In short, we will treat $|\psi\rangle$ as an eigenvector of the group action adjacency matrix with a certain eigenvalue $\lambda$ related to $h$. We run a procedure to estimate $\lambda$ to high precision.
    \item From the estimated eigenvalue(s), we reconstruct a candidate group element $h^*$ that we believe generated the state. (This uses the fact that by running phase estimation for multiple carefully chosen walk durations, one can solve for $h$ as shown in [31].)
    \item We check if $h^* = h$ (the serial number provided). If not, we reject, since the state does not match the serial. If yes, we accept.
\end{enumerate}

@@

\chapter{Verification Algorithm Using Quantum Walks}

>>>>

\usepackage[margin= 1in]{geometry}
\usepackage{article_macros}

\sloppy
\allowdisplaybreaks

@@

\definecolor{citecolor}{HTML}{009933}

\hypersetup{
	unicode = true,
    colorlinks = true,
    citecolor = citecolor,
    filecolor = blue,
    linkcolor = blue,
    urlcolor = blue,
    pdfstartview = {FitH},
}

\title{
  Quantum Walks and Application to Quantum Money\\[1em]
%   \large Supervised by Dr. Jake Doliskani\\[3em]
  \normalsize McMaster University\\
  Department of Computing and Software
}
\author{Ali Mousavi}
\date{April 2025}

@@

\newpage
\textbf{\large Abstract}\\

@@

\medskip\noindent\textbf{This Work – Motivation and Contributions.} 
In this thesis, we adapt Zhandry’s group-action quantum money scheme by replacing its use of the Quantum Fourier Transform with the \emph{Quantum Hartley Transform} (QHT) over finite abelian groups. The motivation behind this substitution is multifold. First, using the Hartley transform causes the banknotes to have \emph{real} amplitudes instead of complex amplitudes. We believe this shift from complex to real quantum states may offer both computational and theoretical advantages. For example, certain mathematical identities hold for any two real orthonormal bases that do not generally hold for complex bases; such differences can affect the properties of quantum states and were a barrier in prior analyses (indeed, an attempted “quantum lightning” construction in \cite{Zhandry2022} failed due to properties of complex phases that do not arise with real amplitudes). Beyond this specific context, our work is a first step toward employing real-valued quantum transforms in quantum cryptography and quantum algorithms. To the best of our knowledge, this is the first significant quantum cryptographic construction that makes essential use of the quantum Hartley transform. We hope that exploring real-amplitude quantum states will inspire further research into new quantum algorithms and optimizations for real-valued transforms.

@@

\medskip\noindent\textbf{Thesis Organization.} 
The remainder of this thesis is organized as follows. In Chapter~\ref{chap:backgound}, we provide background on the basics of quantum computation and the theory of group actions, including the concept of cryptographic group actions which underlie our money scheme. Chapter~\ref{chap:quantum_walks} covers quantum walks, both continuous-time and discrete-time, outlining their differences and importance in quantum algorithms. In Chapter~\ref{chap:simulating}, we apply the quantum walk framework to group actions: we describe how a continuous-time quantum walk on a group action can be simulated efficiently, a crucial step for our serial number extraction algorithm. Chapter~\ref{chap:quantum_money} details the quantum money scheme itself and the role of the Hartley transform in its construction. We explain how the scheme is formulated and discuss the advantages and challenges introduced by using the Hartley transform. Finally, Chapter~\ref{chap:verification} presents the verification algorithm for the Hartley-based quantum money. This chapter shows how we use quantum walks (as developed in Chapter 4) to reliably verify banknotes and compute their serial numbers, thereby addressing the verification issues that arose from the use of real amplitudes.
\vspace{1em}

\chapter{Background of Quantum Computation and Group Actions}\label{chap:backgound}

@@

One example important to this work is the \textbf{quantum Fourier transform} (QFT). For an abelian group $G$ of order $N$, the QFT is the unitary map
\[
|g\rangle \mapsto \frac{1}{\sqrt{N}}\sum_{h \in G} \chi(h,g)\, |h\rangle,
\]
where $\{\chi(h,.)\}$ are the characters (complex exponential homomorphisms) of $G$. When $G = \mathbb{Z}_N$ (the integers mod $N$), the QFT takes $|j\rangle$ to:
\[
\frac{1}{\sqrt{N}}\sum_{k=0}^{N-1} e^{2\pi i j k / N} |k\rangle
\]

Let $G \cong \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$ be a decomposition of $G$. The character $\chi(\bm{a}, x)$ is then given by
\[
\chi(\bm{a}, \bm{x}) = \omega_{N_1}^{a_1x_1} \cdots \omega_{N_k}^{a_kx_k}, \quad \text{where } \omega_M = \exp(2\pi i/ M).
\]

@@

\paragraph{The $\comph$ Algorithm.}
Given a state $\ket{G^{(h)} * x}$, one can recover $h$ using phase kickback. Starting from $\ket{G^{(h)} * x} \ket{0}$, apply:
\begin{itemize}
    \item QFT to second register,
    \item Controlled unitary $\sum_{k \in G} U_k \otimes \ket{k}\bra{k}$,
    \item Inverse QFT to second register.
\end{itemize}
This yields $\ket{G^{(h)} * x} \ket{h}$, extracting $h$ efficiently.

@@

\paragraph{Definition.}
For $G = \mathbb{Z}_N$, define
\[
\qht_N: \ket{a} \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\left(\frac{2\pi a y}{N}\right) \ket{y}, \label{eq:qht_def}
\]
Where $\cas(x) = \cos(x) + \sin(x)$.

For general abelian $G = \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$, the Hartley transform of $f : G \to \mathbb{R}$ is
\begin{equation}
    \qht(f)({\bm{a}}) = \frac{1}{\sqrt{|G|}} \sum_{{\bm{y}} \in G} \cas(2\pi \langle {\bm{a}}, \alpha({\bm{y}}) \rangle) f({\bm{y}}),
\end{equation}
where $\alpha({\bm{y}}) = y_1 / N_1 + \cdots + y_k / N_k$.

\paragraph{Unitary Equivalence.}
Using the identity
\[
\cas(2\pi \langle {\bm{a}}, \alpha({\bm{y}}) \rangle) = \frac{1 - i}{2}\chi({\bm{a}}, {\bm{y}}) + \frac{1 + i}{2}\chi(-{\bm{a}}, {\bm{y}}),
\]
we obtain:
\begin{equation}
\qht_G = \frac{1 - i}{2} \qft_G + \frac{1 + i}{2} \qft_G^*, \label{eq:QHT_QFT}
\end{equation}
showing that $\qht_G^2 = {1}$ and hence the $\qht$ is unitary.

\paragraph{Quantum Formulation.}
The quantum Hartley transform over $G$ maps:
\[
\sum_{{x} \in G} f({x}) \ket{{x}} \mapsto \sum_{{\bm{a}} \in G} \qht_G(f)({\bm{a}}) \ket{{\bm{a}}}.
\]
Because the cas function is real-valued, the resulting amplitudes remain real—a fact that plays a crucial role in real-amplitude quantum protocols.

@@

\chapter{Quantum Walks -- Continuous and Discrete}\label{chap:quantum_walks}

Quantum walks are the quantum-mechanical counterparts of classical random walks on graphs. Instead of a probability distribution over vertices, a quantum walk evolves a superposition of vertex states. Interference between multiple paths leads to markedly different behavior from classical walks, often yielding faster spreading and algorithmic speedups~\cite{Aharonov2001}.

In particular, both major models of quantum walks --- continuous-time and discrete-time --- achieve \emph{ballistic spreading} (standard deviation growing linearly in time) in contrast to the diffusive $O(\sqrt{t})$ spread of classical random walks~\cite{Kempe2003}. Quantum walks have become a central framework in quantum computing, underpinning various algorithms and even universal quantum computation~\cite{Childs2009,Ambainis2003}.

@@

Continuous-time quantum walks have been employed to achieve algorithmic speedups. A famous example is the \emph{glued trees traversal problem} by Childs \cite{Childs2003}, where a CTQW on a graph formed by two randomly "glued" binary trees can find the unique exit vertex exponentially faster than any classical algorithm. This result gave the first exponential separation between quantum and classical query complexity in a black-box setting.

Another application is \emph{spatial search} on graphs. Childs and Goldstone \cite{ChildsGoldstone2004} showed that a CTQW can find a marked vertex on a $d$-dimensional lattice faster than classical diffusion. In the limit $d \to \infty$, the CTQW dynamics converge to the continuous Dirac equation.

@@

In 2004, Szegedy \cite{Szegedy2004} proposed a different formulation of DTQW based on classical Markov chains. Here, no coin register is needed. Instead, the walk is defined on $\mathbb{C}^N \otimes \mathbb{C}^N$, with basis states $\ket{j,k}$ corresponding to directed edges.

@@

Both DTQW and CTQW have been used to achieve quadratic speedups in search problems. In many cases, the two models can simulate each other. DTQWs can even achieve \textbf{universal quantum computation}: Childs\cite{Childs2009} constructed a coined DTQW that simulates an arbitrary quantum circuit.

@@

    \item \textbf{Infinite Tree (Free Group):} The free group on two generators $a$ and $b$ has a Cayley graph that is an infinite 4-regular tree. Each vertex corresponds to a word in $\{a^{\pm1}, b^{\pm1}\}$, and edges represent multiplication by $a$ or $b$.

@@

To see this, we have

        \begin{align}
            A\ket{\hat{a}} 
            &= A\frac{1}{\sqrt{|G|}}\sum_{y \in G}\chi(a,y)\ket{y} \nonumber \\
            &= \frac{1}{\sqrt{|G|}}\sum_{y \in G}\chi(a,y)A\ket{y} \nonumber \\
            &= \frac{1}{\sqrt{|G|}}\sum_{y \in G}\chi(a,y)\sum_{q \in Q}\ket{qy} \nonumber
            \end{align}

            \noindent Consider $\beta = qy$. Then the above equation becomes:

            \begin{align}
            \frac{1}{\sqrt{|G|}}\sum_{q \in Q} \sum_{y \in G} \chi(a,y)\ket{qy} 
            &= \frac{1}{\sqrt{|G|}}\sum_{q \in Q} \chi(a,q) \sum_{\beta \in G} \chi(a,\beta)\ket{\beta} \nonumber \\
            &= \sum_{q \in Q} \chi(a,q)\ket{\hat{a}} \nonumber \\
            &= \lambda_a\ket{\hat{a}} \label{eq:group_action_walk}
            \end{align}

@@

In particular, for a continuous-time quantum walk starting at the identity element:
\[
\ket{\psi(t)} = e^{-iAt} \ket{e},
\]
the resulting state remains a \emph{class function}, i.e., the amplitude $\braket{g}{\psi(t)}$ depends only on structural properties of $g$ (such as Hamming weight or conjugacy class).

@@

\begin{align}
    A\ket{G^{(h)} * x} 
    &= \frac{1}{\sqrt{|G|}}\sum_{g \in G} \chi(g, h)\, A\ket{g * x} \nonumber \\
    &= \frac{1}{\sqrt{|G|}}\sum_{g \in G} \chi(g, h)\sum_{q \in Q} \ket{q * (g * x)} \nonumber
    \end{align}

    \noindent Consider $\beta = qg$. Then the above equation becomes:

    \begin{align}
    \frac{1}{\sqrt{|G|}}\sum_{q \in Q} \sum_{\beta \in G} \chi(\beta, h)\chi(q, h)\ket{\beta * x}
    &= \sum_{q \in Q} \chi(q, h)\ket{G^{(h)} * x} \nonumber \\
    &= \lambda_h \ket{G^{(h)} * x} \label{eq:group_action_symmetrized}
    \end{align}

@@

\chapter{Simulating Group Action Quantum Walks}\label{chap:simulating}

From this point onward, we assume that the generating set \( Q \) is small, specifically \( \abs{Q} = \operatorname{poly}(\log \abs{G}) \). In the first construction, where the vertex set is \( X = G \), the continuous-time walk \( e^{-iAt} \) can be implemented efficiently even for values of \( t \) that are exponentially large in \( \log \abs{G} \). This efficiency stems from the identity:
\begin{equation}
    \label{eq:adjmat-fb}
    e^{-iAt} = \qft_G \left( \sum_{a \in G} e^{-i\lambda_a t} \ket{a} \bra{a} \right) \qft_G^*,
\end{equation}
where both \( \qft_G \) and its inverse can be implemented in \( \operatorname{poly}(\log \abs{G}) \) time. Moreover, the diagonal unitary \( \ket{a} \mapsto e^{-i\lambda_a t} \ket{a} \) is efficiently realizable since the eigenvalues \( \lambda_a \) can be computed classically to arbitrary precision in polynomial time with respect to \( \log \abs{G} \).

@@

Based on these states, we define the isometry \( T : \C^{2N} \to \mathbb{C}^{2N} \otimes \mathbb{C}^{2N} \) as follows:
\begin{equation}
    \label{eq:walk-isom}
    T := \sum_{j = 0}^{N - 1} \sum_{b \in \{0, 1\}} (\ket{j}\bra{j} \otimes \ket{b}\bra{b}) \otimes \ket{\phi_{jb}}.
\end{equation}

The discrete-time quantum walk operator is then defined as
\[
W = iS(2TT^* - \mathds{1}),
\]
where \( S \) is the swap operator acting as
\[
S \ket{j_1} \ket{b_1} \ket{j_2} \ket{b_2} = \ket{j_2} \ket{b_2} \ket{j_1} \ket{b_1},
\]
for all \( 0 \le j_1, j_2 < N \) and \( b_1, b_2 \in \{0, 1\} \).

@@

\section{Simulating group action quantum walks.}
Assume \( t = \operatorname{poly}(\log \abs{G}) \). We show that the continuous-time walk \( W = e^{-iAt} \) can be efficiently simulated using the discrete-time quantum walk framework developed in~\cite{childs2010relationship}. Recall the isometry \( T \) defined in Equation~\eqref{eq:walk-isom}, where the states \( \ket{\phi_b} \) are as given in Equation~\eqref{eq:walk-sim}. To carry out the simulation, it suffices to show that the isometry \( T \), its adjoint \( T^* \), and the discrete-time walk operator
\[
W = iS(2TT^* - \mathds{1})
\]
can all be efficiently implemented in the group-action setting.

@@

Because \( Q \subset G \) is a small set (i.e., \( \abs{Q} = \operatorname{poly}(\log \abs{G}) \)), we can construct an efficient unitary \( V_Q : \mathbb{C}^Q \to \mathbb{C}^Q \) such that
\[
V_Q \ket{0} = \frac{1}{\sqrt{\abs{Q}}} \sum_{q \in Q} \ket{q}.
\]
Applying \( V_Q \otimes \mathds{1} \) to the state \( \ket{0} \ket{y, 0} \ket{0, 0} \) produces:
\[
\frac{1}{\sqrt{\abs{Q}}} \sum_{q \in Q} \ket{q} \ket{y, 0} \ket{0, 0}.
\]

@@

To implement \( U_0^* : \ket{0} \ket{y, 0} \ket{\phi_{y0}} \mapsto \ket{0} \ket{y, 0} \ket{0, 0} \), we sequentially apply \( V_2^* \), followed by \( V_1^* \), and finally \( V_Q^* \otimes \mathds{1} \). Since each of these unitaries can be implemented efficiently, it follows that \( U_0^* \) is also efficient.

@@

Next, we demonstrate that the walk unitary
\[
W = iS(2TT^* - \mathds{1})
\]
can be applied efficiently. This reduces to implementing the reflection \( 2TT^* - \mathds{1} \), which we analyze as follows:
\begin{align*}
2\ket{0}\bra{0} \otimes TT^* - \mathds{1}
&= 2\ket{0}\bra{0} \otimes \sum_{y \in X} \sum_{b \in \{0,1\}} \ket{y,b} \ket{\phi_{yb}} \bra{y,b} \bra{\phi_{yb}} - \mathds{1} \\
&= 2\sum_{y \in X} \sum_{b \in \{0,1\}} \ket{0} \ket{y,b} \ket{\phi_{yb}} \bra{0} \bra{y,b} \bra{\phi_{yb}} - \mathds{1} \\
&= U_T \left( 2 \sum_{y \in X} \sum_{b \in \{0,1\}} \ket{0} \ket{y,b} \ket{0,0} \bra{0} \bra{y,b} \bra{0,0} - \mathds{1} \right) U_T^* \\
&= U_T \left( 2\ket{0}\bra{0} \otimes \mathds{1}_{X,b} \otimes \ket{0,0}\bra{0,0} - \mathds{1} \right) U_T^*.
\end{align*}

Since \( U_T \), \( U_T^* \), and the operator \( 2\ket{0}\bra{0} \otimes \mathds{1}_{X,b} \otimes \ket{0,0}\bra{0,0} - \mathds{1} \) can all be implemented efficiently, it follows that the reflection \( 2\ket{0}\bra{0} \otimes TT^* - \mathds{1} \) is also efficient.

Finally, for any input state \( \ket{\psi} \), the reflection acts as:
\[
\left(2\ket{0}\bra{0} \otimes TT^* - \mathds{1} \right) \ket{0} \ket{\psi} = \ket{0} \left(2TT^* - \mathds{1} \right) \ket{\psi},
\]
confirming that the full walk operator \( W \) can be implemented efficiently in the group-action setting.

@@

\chapter{Quantum Money and Hartley Transform}\label{chap:quantum_money}

@@

\section{Fourier-Based Quantum Money from Group Actions}
Zhandry’s quantum money scheme \cite{Zhandry2024} is built on an abelian group action $G \curvearrowright X$. We have a family of group actions parameterized by the security parameter $n$ (so the size of $G$ grows with $n$). For simplicity, assume $|G| = N$. The scheme also fixes a particular element $x_0 \in X$ that everyone knows (a public base point in $X$).

@@

To address this, our verification algorithm takes a more sophisticated approach:
\begin{enumerate}
    \item Given $(h, |\psi\rangle)$, it first uses the \textbf{quantum walk-based phase estimation} method (to be detailed in Chapter 6) to \emph{compute the eigenphase corresponding to $h$}. In short, we will treat $|\psi\rangle$ as an eigenvector of the group action adjacency matrix with a certain eigenvalue $\lambda$ related to $h$. We run a procedure to estimate $\lambda$ to high precision.
    \item From the estimated eigenvalue(s), we reconstruct a candidate group element $h^*$ that we believe generated the state. (This uses the fact that by running phase estimation for multiple carefully chosen walk durations, one can solve for $h$ as shown in \cite{Zhandry2022}.)
    \item We check if $h^* = h$ (the serial number provided). If not, we reject, since the state does not match the serial. If yes, we accept.
\end{enumerate}

@@

\chapter{Verification Algorithm Using Quantum Walks}\label{chap:verification}

The Hartley-based quantum money scheme introduced in the previous chapter requires a new verification procedure to address the issues that arise from using real amplitude states. In this chapter, we describe and analyze the verification algorithm, which leverages continuous-time quantum walks on the group action graph (and their efficient simulation from Chapter~\ref{chap:simulating}) to extract the information needed to authenticate a banknote.
